Autonomous Oceanographic Sampling Designs using Excursion Sets for Multivariate Gaussian random fields

### 2020-06-04 main_new.tex
Updates from Overleaf

With this in mind, we have designed AUV sampling strategies for
reducing the spatially-integrated Bernoulli variance (IBV) in the ES
indicator variables. Recent work on ESs and associated excursion
probabilities (EPs) connected to spatial statistics include
\cite{french2013spatio,bolin2015excursion,french2016credible}. Our
focus is on uncertainty characterization and how this is influenced by
sampling. In this sense, our work is similar to
\cite{bect2012,chevalier2014fast,azzimonti2016quantifying} who
describe analytical results for the IBV for univariate processes.

@@

\kc{Typically, I have been told in the past, it is bad form to start a section and
  then a subsection right away. I sort of see it, so, it might be good to
  provide a 1-2 sentence introduction here perhaps?}

\subsection{Excursion sets and Integrated Bernoulli variance}
\label{ES}

@@

In order to formalize the problem in mathematical terms, let us
consider a multivariate (i.e., vector-valued) Gaussian random field
$(\gp[\x])_{\x \in \mathcal{M}}$, where $\mathcal{M}$ is an arbitrary
set (typically in geostatistical studies, a subset of interest of the
ambient Euclidean space or of a sphere therein \kc{is this obvious to
  the readership why a sphere? It isn't for me certainly.}) and the
dimensionality of the output $\no \geq 1$ is arbitrary \kc{Is the
  definition of such a 'p' and its dimensionality, obvious? This to me
  springs out of nowhere.}. Assume furthermore that the set of interest
is characterized in terms of a subset of values of the variables, say
$\T \subset \R^{\no}$, and denote the corresponding set of index
points by
$$
\es=\gp^{-1}(\T)=\{\x \in \mathcal{M}: \gp[\x] \in \T\}.
$$
Our goal here is to develop approaches to quantify and improve the
characterization of uncertainties on such sets \kc{$\es$?} in the
context of multivariate random fields.

@@

For statistical generality, we will however keep the abstract notation
with $\T$ and general $\es$ \kc{Again, previous phrase doesn't parse.}
for most of the derivations. Note that assuming that $\gp$ has
continuous trajectories (almost surely) and $T$ to be closed, then
$\es$ can be cast as a Random Closed Set
\citep{Molchanov2005,adler2009random}. %In what follows, we will express various quantities related to $\es$ in terms of $\gp$'s moments both for general and orthant $T$'s.
Let us in particular consider a locally finite (Borel) measure $\mes$
on $D$ and investigate the probability distribution of $\mes(\es)$
through its moments.

@@

Proposition~\ref{propo1} enables one to calculate centered moments of $\mes(\es)$. 
The excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$
is probably the most relevant of these for our purposes. We
have 
\begin{equation*}
\begin{split}
\emv
&=\int_{D^2} \mathbb{P}\left(
\gp[u]\in T, \gp[v]\in T \right) 
d\mes^{\otimes}(u, v)\\
&-\left( \int_{D} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
\end{split}
\end{equation*} 
which evaluates in the excursion/sojourn case where
$\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ to
\begin{equation*} \begin{split}
  \emv
%\operatorname{Var}[\mes(\es)]
&=\int_{D^2} 
\varPhi_{2\no}
\left(
(\bt, \bt); \bmu((u,v)), 
\bK((u,v),(u,v))
\right) 
\
\mathrm{d}\mes^{\otimes} %\mes 
%\productMeasure 
(u,v)\\
&-\left( \int_{D} \varPhi_{\no}\left(\bt;\bmu(u), \bK(u)\right) d\mes(u) \right)^2.
\end{split}
\end{equation*}
This quantity requires  an integral over $D^2$. In
contrast, the Integrated Bernoulli Variance (IBV) of \cite{Bect.etal}
involves solely an integral on $\mathcal{M}$ and can be expanded in
our present setting as follows 
\begin{equation}\label{IBV}
\begin{split}
\operatorname{IBV} %(\es) %;\mes)
&=\int_{\mathcal{M}}
\mathbb{P}\left(\gp[\uu]\in T\right)(1-\mathbb{P}\left(\gp[\uu]\in T\right))
d\mes(u) \\
&=\int_{\mathcal{M}}
\varPhi_{\no}\left(\bt;\bmu(\uu), \bK(\uu)\right) 
-\left(\varPhi_{\no}\left(\bt;\bmu(\uu), \bK(\uu)\right) \right)^2
\mathrm{d}\mes(u). 
\end{split}
\end{equation}

\subsection{Expected Integrated Bernoulli Variance and Sampling Designs}

@@

We let $d \in \mathcal{D}$ denote a sampling design, with $q$ sampling 
locations $\bu_d=\left(u_d^{(1)},\ldots, u_d^{(q)}\right)$. The
measurement model is defined by 
\begin{equation}\label{eq:meas}
    \bY_d=\bG_d \bz(\bu_d) + \epsilon(\bu_d), \hspace{3mm} \epsilon(\bu_d) \sim N(0,\bR_d).
\end{equation}
Here, $\bz(\bu_d)$ is the size $pq$ vector of GP variables at the
sampling locations, $\bG_d$ is the design matrix which in principle
allows for observing subsets of the vector-valued field at some of the
$q$ locations (i.e. the matrix has fewer rows than columns). If the
additive error has independent elements, the covariance matrix $\bR_d$
is diagonal. In our setting temperature and salinity measurements are
extracted from different sensors, so the independence assumption
between variables is not unrealistic, but this is not necessarily the
case for other kinds of measurements on the AUV. It is further common
in geostatistics to assume spatially independent measurement errors,
and we also do so. But our methodology for constructing sampling
designs does not rely on these independence assumptions.

For a given design $d$ the evaluation is based on the expected
information contained in the data to influence the criterion.   
Considering the EIBV criteria, we have
\begin{equation}\label{eq:eibv}
    \mbox{EIBV} (d)= E_{\bY(\bu_d)} \int_{M}
\mathbb{P}\left(\gp[\uu]\in T|\by(\bu_d) \right)(1-\mathbb{P}\left(\gp[\uu]\in T |\by(\bu_d)\right))
d\mes(u),
\end{equation}
where the expectation is over the planned data.  The optimal design
for an AUV survey is then 
\begin{equation}\label{crit}
    d^* = \mbox{argmin}_{d \in \mathcal{D}} \mbox{EIBV}(d).
\end{equation}
For another criterion using random moments of random sets similar
expectations would be required, but as we will see, the EIBV has a
closed-form expression which is relatively fast to compute and hence
useful for onboard design selection.

Assuming now that $q$ observations are available, one wishes to
predict $\gp[\bu]$, the multivariate Kriging equations holds with size
$p$ vector of conditional means \kc{This last sentence doesn't parse.
  Are they two separate statements?}:  
\begin{equation}\label{eq:cokrig}
\bbm(\x)=\bmu(\x)+\blambda(\x,\bu_d)^T \left( \by_d-\bG_d \bmu(\bu_d)\right),
\end{equation}
where $\bmu$ is size $p$ unconditional mean function of $Z$, $\by_d$
is the realized vector of observed responses, and
$\blambda(\x,\bu_d)=\bK^{-1}_y(\bu_d,\bu_d) \bk(\bu_d, \bu)$. Here,
the data covariance is
$K_y=\mbox{Cov}(\bY_d,\bY_d)=\bG_d\bK(\bu_d,\bu_d)\bG_d^T+\bR_d$ and
the cross-covariance is $\mbox{Cov}(\bY_d,Z(\x)) \bk(\bu_d, \x)=$,
\kc{There is a dangling '='. Is this legit?} assuming the measurement
error is independent of the GP.  The Kriging 
(cross-)covariance function can also be expressed in the same vein via
a $p \times p$ matrix 
\begin{equation}\label{eq:krig:cov}
\bSigma(\x,\x')=\bK(\x,\x')-\blambda(\x,\bu_d)^T \bK_y(\bu_d, \bu_d) \lambda(\x,\bu_d).
\end{equation}

Equations \eqref{eq:cokrig} and \eqref{eq:krig:cov} are used to
provide informative, updated statements about sets of interest. For
our case with temperature and salinity, the conditional EPs, given
measurements $\by_d$, can then be computed from
\begin{eqnarray}\label{eq:post_ep}
 p(\x;\by) &=& P(Z_1(\x) \leq t_1, Z_{2}(\x) \leq t_2 |\by(\bu_d)),  \nonumber \\
 &=& \Phi_2 ((t_1,t_2)^T-\bmu(\x)-\blambda(\x,\bu_d)^T \left( \by(\bu_d)-\bG_d \bmu(\bu_d)\right);0_2, \bSigma(\bu,\bu)). 
\end{eqnarray}

A critical element in the derivation of a closed-form for the EIBV is
that the conditional mean in Eq. \eqref{eq:cokrig} is a linear
(affine) function of the data $\by_d$ and the covariance is not a
function of the data. This in turn means that the probabilities in
\eqref{eq:post_ep} involve inequality statements for linear
combinations of Gaussian variables. Related closed-form solutions have
been noted in similar contexts \citep{bhattacharjya2013value,
  chevalier2014fast,stroh}, but not generalized to our situation with
random sets for vector-valued GPs. Notably, the mean in equation
\eqref{eq:cokrig}, is in the form of $\ba+\bB \bV$, where
$\bV \sim N(0_q, \bK_y (\bu_d,\bu_d))$,
$\ba=(t_1,t_2)^t-\bmu(\bu)-\blambda(\bu,\bu_d)^T \bG_d \bmu(\bu_d)$
and $\bB=\blambda(\bu,\bu_d)^T$. The following then applies to give a
closed-form solution to EIBV calculation in this case.

\begin{propo}
	\label{propo2}
	Let $p, q, h \geq 1$, $a \in \R^p$, $B \in \R^{p\times q}$,
	and $\covN, \covV$ be two covariance matrices in
	$\R^{p\times p}$ and $\R^{q\times q}$, respectively.
	Then, for $V \sim \mathcal{N}_{q}(0_q, \covV)$,
	\begin{equation*}
	\mathbb{E}\left[ \varPhi_{p}\left( a + BV; 0_p,\covN \right)^h \right]
	=
	\varPhi_{ph}
	\left(
	\bm{a}
	;0_{ph},
	\bm{\Sigma}
	\right),
	\end{equation*}
	where the vector $\bm{a} \in \R^{p h}$ is defined as 
	$\bm{a} := 1_h\otimes a = 
	\left(a, \dots , a
	\right)'$
	and the $p h\times p h$ covariance matrix is given by
	$\bm{\Sigma} := 
	1_h 1_h'\otimes B\covV B' + I_h\otimes \covN$.
\end{propo}
The proof of Proposition \ref{propo2} is provided in the Appendix.

In block-wise representation, the following relation holds for the
covariance in Proposition \ref{propo2}: 
\begin{align*}
	\bm{\Sigma}&=
	\begin{pmatrix}
	\covN & &\\
	& \ddots &\\
	&   & \covN
	\end{pmatrix}
	+
	\begin{pmatrix}
	B\covV B' & \dots & B\covV  B'\\
	\vdots & & \vdots\\
	B\covV B' & \dots & B\covV B'\\
	\end{pmatrix}.
\end{align*}
The calculation of the multivariate Gaussian cumulative distribution
is done effectively using code such as that of
\cite{genz2009computation}.

@@

\subsection{An Example of EIBV using Temperature and Salinity}

@@

Table \ref{tab:sim_rhoab} (bottom two rows) shows results of expected
Bernoulli variance calculations. This is presented for a design
gathering both data types, and for a design with
temperature measurements alone. Both data are modeled by $(Y_1,Y_2)^t=(Z_1,Z_2)^t+N(0,0.5^2\bI)$, while
$Y_2=Z_2+N(0,0.5^2)$ when only temperature is measured.  For this
illustration, Table \ref{tab:sim_rhoab} shows that the expected
Bernoulli variance gets lower with larger standard deviations
$\sigma_1$ and $\sigma_2$ (right columns). The reduction of Bernoulli
variance is largest for the cases with high correlation
$\gamma$. Albeit smaller, there is also uncertainty reduction when
only temperature is measured (bottom row), especially when temperature
and salinity are highly correlated. When correlation is low
($\gamma=0.2$), there is little information about salinity in the
temperature data, and therefore less uncertainty reduction. In an
application with fresh cold water from a river source, the temperature
and salinity variables will not only be interdependent, but will also
likely show dependence in the spatial dimension. This in turn will
impact the design criteria when we evaluate the information measure by
integrating over several locations (Section \ref{sec:simulations}).

@@

An adaptive AUV survey is split into many stages. The selection on where to sample is made with restriction to the waypoint graph and defined by the nearest grid nodes in the domain $\mathcal{M}_g$.

Let $d^{j,s}$ denote design
number $j$ at stage $s$ of a sequential survey. If this design is
selected, data $\by_{d^{j,s}}$ will be gathered. 
We further let
$\mathcal{Y}^{s-1} =\{(d^{j,1}, \by_{d^{j,1}}),\ldots, (d^{j,s}, \by_{d^{j,s}}) \}$ denote
the set of all data gathered until stage $s-1$ according to selected design $d^{j,1},\ldots,d^{j,s}$. 
This defines the AUV position at the stage $s$. 
The data further defines the conditional GP  with mean $\bbm^{s-1}(\x)=E(\gp[u]|\mathcal{Y}^{s-1})$ from
Eq. (\ref{eq:cokrig})
and (cross) covariance matrix $\bSigma^{s-1}=\mbox{Var}(\gp[u]|\mathcal{Y}^{s-1})$ from Eq. (\ref{eq:krig:cov}).

@@

The simplest heuristic for adaptive sampling is to choose the next
sampling location based on current probabilities. One reduces the largest levels of uncertainty by selecting the design $j$ with $P(Z(\x_j) \in T)$ closest to $0.5$.
For our situation with temperature and salinity this {\it{naive}} strategy then
selects the design according to:
\begin{eqnarray}\label{critNaive}
    d^{*,s} &=& \mbox{argmin}_{j} |P(Z_1(\x_j) \leq t_1, Z_2(\x_j) \leq t_2 | \mathcal{Y}^{s-1})-0.5|. \nonumber
\end{eqnarray}

@@

Based on the currently available data $\mathcal{Y}^{s-1}$, we
fit an updated GP model. Focusing on the EIBV criterion, at the next stage $s$, the selected design is:
\begin{eqnarray}\label{critSEQ}
    d^{*,s} &=& \mbox{argmin}_{j} \left\{ \sum_{\x \in \mathcal{M}_g} E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left( p(\x;\mathcal{Y}^{j,s})\left( 1-p(\x;\mathcal{Y}^{j,s})\right) \right) \Delta, \right\} \\
    p(\x;\mathcal{Y}^{j,s}) &=& P(Z_u \in T |\by_{d^{j,s}},\mathcal{Y}^{s-1}). \nonumber
\end{eqnarray}

Note that this strategy gives a sequential conditional version of
expression (\ref{eq:eibv}). Now $\mathcal{Y}^{s-1}$ is available, and
the expectation is conditional to this data. A similar closed-form calculation
for EIBV is hence applicable in Eq. (\ref{critSEQ}) using the
updated GP model from stage $s-1$. Once the data is collected for the
best design at stage $s$, the GP model is updated again. The mean $\bbm^{s}$ and
covariance matrix $\bSigma^{s}$ are used to compute the next design at
stage $s+1$, and so on.

@@

We now extend the myopic strategy to a look-ahead strategy, which is
optimal when one can gather only two more stages of measurements. In
addition to the next stage of measurements $\by^{j,s}$, this
look-ahead strategy, anticipates the subsequent design $j_2$ with
data $\by^{j_2,s+1}$ when choosing the current design $d^{j,s}$.  The
selected design is:
\begin{eqnarray}\label{critLA}
    d^{*,s} &=& \mbox{argmin}_{j} \left\{ E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left[ \mbox{argmin}_{j_2}  V_{j_2} \right] \right\}, \nonumber \\
V_{j_2} & = & \sum_{\x \in \mathcal{M}_g} E_{\by_{d^{j_2,s+1}}|\mathcal{Y}^{j,s}} \left( p(\x;\mathcal{Y}^{j,s+1})\left( 1-p(\x;\mathcal{Y}^{j,s+1})\right) \right) \Delta, \nonumber \\
    p(\x;\mathcal{Y}^{j,s+1}) &=& P(Z(\x) \in T |\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}). \nonumber
\end{eqnarray}
Here, $\mathcal{Y}^{j,s}=\{\by_{d^{j,s}},\mathcal{Y}^{s-1}\}$ and
$\mathcal{Y}^{j,s+1}=\{\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}\}$ represent
the sets of data variables, and the expectations are calculated with respect to
the conditional densities given data at that stage, meaning we have two levels of intermixed minimizations and integrals.
We solve the
first expectation by Monte Carlo sampling of data $\by_{d^{j,s}}$ from its
conditional distribution. For each of these data samples, the second expectation
is solved using the closed-form expressions for EIBV from
Section \ref{sec:ESEP}.

Even though the strategy looks at two sampling stages, it is only used to find
the current best design. When data are collected, the GP model is
updated, and the mean $\bbm^{s}$ and covariance matrix $\bSigma^{s}$ are
used to compute the next design at stage $s+1$, now anticipating what
stage $s+2$ could offer, and so on.
This look-ahead approach is much more computationally demanding than
the myopic strategy, so for practical implementation, we prune paths
in the evaluation of Eq. (\ref{critLA}). This means that we use the myopic
strategy to rank the three best designs following the first stage alone, and
for each of these preferred designs, we undertake the look-ahead calculations.

@@

      \caption{Simulation results for the final mean explained variance $\bR^{2}=100*(1-(\bSigma_{posterior}/\bSigma_{initial}))$).}

>>>>

With this in mind, we have designed AUV sampling strategies for
reducing the spatially-integrated Bernoulli variance (IBV) in the ES
indicator variables. 
ES of random fields are thoroughly presented in \citep{adler2009random}. Recent work on ESs and associated excursion
probabilities (EPs) connected to spatial statistics include
\cite{french2013spatio,bolin2015excursion,french2016credible}. Our
focus is on uncertainty characterization and how this is influenced by
sampling. In this sense, our work is similar to
\cite{bect2012,chevalier2014fast,azzimonti2016quantifying} who
describe analytical results for the IBV for univariate processes.

@@

The definitions of the sets of interest are introduced in Section \ref{sec:ES}, while the closed-form expressions for the expected uncertainty reduction are derived in Section \ref{sec:EIBV}. Section \ref{Sec:UnivarEx} provides an instructive bivariate example relevant for sampling in the temperature and salinity application.

\subsection{Excursion sets and Integrated Bernoulli variance}
\label{sec:ES}

@@

In order to formalize the problem in mathematical terms, let us
consider a multivariate (i.e., vector-valued) Gaussian random field
$(\gp[\x])_{\x \in \mathcal{M}}$, where $\mathcal{M}$ is an arbitrary
set (typically in geostatistical studies, a geographical subset of interest) and the
dimensionality of the output $\no \geq 1$ is arbitrary \kc{Is the
  definition of such a 'p' and its dimensionality, obvious? This to me
  springs out of nowhere.}. Assume furthermore that the set of interest
is characterized in terms of a subset of values of the variables, say
$\T \subset \R^{\no}$, and denote the corresponding set of index
points by
$$
\es=\gp^{-1}(\T)=\{\x \in \mathcal{M}: \gp[\x] \in \T\}.
$$
Our goal here is to develop approaches to quantify and improve the
characterization of uncertainties on such sets \kc{$\es$?} in the
context of multivariate random fields.

@@

For statistical generality, we will however keep the abstract notation
with $\T$ and general $\es$ \kc{Again, previous phrase doesn't parse.}
for most of the derivations. Note that assuming that $\gp$ has
continuous trajectories (almost surely) and $T$ to be closed, then
$\es$ can be cast as a Random Closed Set
\citep{Molchanov2005}. %In what follows, we will express various quantities related to $\es$ in terms of $\gp$'s moments both for general and orthant $T$'s.
Let us in particular consider a locally finite (Borel) measure $\mes$
on $D$ and investigate the probability distribution of $\mes(\es)$
through its moments.

@@

Proposition~\ref{propo1} enables one to calculate centered moments of $\mes(\es)$. 
The excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$
is probably the most relevant of these for our purposes. We
have 
\begin{equation*}
\begin{split}
\emv
&=\int_{\mathcal{M}^2} \mathbb{P}\left(
\gp[u]\in T, \gp[v]\in T \right) 
d\mes^{\otimes}(u, v)\\
&-\left( \int_{\mathcal{M}} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
\end{split}
\end{equation*} 
which evaluates in the excursion/sojourn case where
$\T=(-\infty, t_1] \times \dots \times (-\infty, t_{\no}]$ to
\begin{equation*} \begin{split}
  \emv
%\operatorname{Var}[\mes(\es)]
&=\int_{\mathcal{M}^2} 
\varPhi_{2\no}
\left(
(\bt, \bt); \bmu((u,v)), 
K((u,v),(u,v))
\right) 
\
\mathrm{d}\mes^{\otimes} %\mes 
%\productMeasure 
(u,v)\\
&-\left( \int_{\mathcal{M}} \varPhi_{\no}\left(\bt;\bmu(u), K(u)\right) d\mes(u) \right)^2.
\end{split}
\end{equation*}
This quantity requires  an integral over $D^2$. In
contrast, the Integrated Bernoulli Variance (IBV) of \cite{Bect.etal}
involves solely an integral on $\mathcal{M}$ and can be expanded in
our present setting as follows 
\begin{equation}\label{IBV}
\begin{split}
\operatorname{IBV} %(\es) %;\mes)
&=\int_{\mathcal{M}}
\mathbb{P}\left(\gp[\uu]\in T\right)(1-\mathbb{P}\left(\gp[\uu]\in T\right))
d\mes(u) \\
&=\int_{\mathcal{M}}
\varPhi_{\no}\left(\bt;\bmu(\uu), K(\uu)\right) 
-\left(\varPhi_{\no}\left(\bt;\bmu(\uu), K(\uu)\right) \right)^2
\mathrm{d}\mes(u). 
\end{split}
\end{equation}

\subsection{Expected Integrated Bernoulli Variance and Sampling Designs}
\label{sec:EIBV}

@@

We let $d \in \mathcal{D}$ denote a concatenation of sampling designs with $q$ sampling locations $\bu_d=\left(u_d^{(1)},\ldots, u_d^{(q)}\right)$. The
measurement model is defined by 
\begin{equation}\label{eq:meas}
    \bY_d=G_d \bz(\bu_d) + \epsilon, \hspace{3mm} \epsilon \sim N(0,R_d).
\end{equation}
Here, $\bz(\bu_d)$ is the size $pq$ vector of GP variables at the
sampling locations, $G_d$ is the design matrix which in principle
allows for observing subsets of the vector-valued field at some of the
$q$ locations (i.e. the matrix has fewer rows than columns). If the
additive error has independent elements, the covariance matrix $R_d$
is diagonal. In our setting temperature and salinity measurements are
extracted from different sensors, so the independence assumption
between variables is not unrealistic, but this is not necessarily the
case for other kinds of measurements on the AUV. It is further common
in geostatistics to assume spatially independent measurement errors,
and we also do so. But our methodology for constructing sampling
designs does not rely on these independence assumptions.

For a given design $d$ the evaluation is based on the expected
information contained in the data to influence the criterion.   
Considering the EIBV criteria, we have
\begin{eqnarray}\label{eq:eibv}
    \mbox{EIBV} (d) &=& E_{\bY_d} \left( \int_{\mathcal{M}}
p(\x;\by_d)(1-p(\x;\by_d))
d\mes(u) \right), \nonumber \\
p(\x;\by_d) &=& \mathbb{P}\left(\gp[\uu]\in T|\bY_d=\by_d \right)
\end{eqnarray}
where the expectation is over the planned data.  The optimal design
for an AUV survey is then 
\begin{equation}\label{crit}
    d^* = \mbox{argmin}_{d \in \mathcal{D}} \mbox{EIBV}(d).
\end{equation}
For another criterion using random moments of random sets similar
expectations would be required, but as we will see, the EIBV has a
closed-form expression which is relatively fast to compute and hence
useful for onboard design selection.

Assuming now that $q$ observations are available, one wishes to
predict $\gp[\x]$. The multivariate Kriging equations holds with size
$p$ vector of conditional means:  
\begin{equation}\label{eq:cokrig}
\bbm(\x)=\bmu(\x)+\lambda(\x,\bu_d)^T \left( \by_d-G_d \bmu(\bu_d)\right),
\end{equation}
where $\bmu$ is size $p$ unconditional mean function of $Z$, $\by_d$
is the realized vector of observed responses, and
$\lambda(\x,\bu_d)=K^{-1}_y(\bu_d,\bu_d) k(\bu_d, \bu)$. Here,
the data covariance is
$K_y=\mbox{Cov}(\bY_d,\bY_d)=G_d K(\bu_d,\bu_d)G_d^T+R_d$ while
the cross-covariance is $\mbox{Cov}(\bY_d,Z(\x))=G_d k(\bu_d, \x)$, assuming the measurement
error is independent of the random fields.  The Kriging (cross-)covariance function can also be expressed in the same vein via
a $p \times p$ matrix 
\begin{equation}\label{eq:krig:cov}
\Sigma(\x,\x')=K(\x,\x')-\lambda(\x,\bu_d)^T K_y(\bu_d, \bu_d) \lambda(\x,\bu_d).
\end{equation}

Eq. \eqref{eq:cokrig} and \eqref{eq:krig:cov} are used to
provide informative, updated statements about sets of interest. For
our case with temperature and salinity, the conditional EPs, given
measurements $\by_d$, can then be computed from
\begin{eqnarray}\label{eq:post_ep}
 p(\x;\by_d) &=& P(Z_1(\x) \leq t_1, Z_{2}(\x) \leq t_2 |\bY_d=\by_d),  \nonumber \\
 &=& \Phi_2 ((t_1,t_2)^T-\bbm(\x);0_2, \Sigma(\x,\x)). 
\end{eqnarray}

A critical element in the derivation of a closed-form for the EIBV is
that the conditional mean in Eq. \eqref{eq:cokrig} is a linear
(affine) function of the data $\by_d$ and the covariance is not a
function of the data. This in turn means that the probabilities in Eq.
\eqref{eq:post_ep} involve inequality statements for linear
combinations of Gaussian variables. Related closed-form solutions have
been noted in similar contexts \citep{bhattacharjya2013value,
  chevalier2014fast,stroh}, but not generalized to our situation with
random sets for vector-valued GPs. Notably, the mean in equation
\eqref{eq:cokrig}, is in the form of $\ba+B V$, where
$V \sim N(0_q, K_y (\bu_d,\bu_d))$,
$\ba=(t_1,t_2)^t-\bmu(\x)+\lambda(\x,\bu_d)^T G_d \bmu(\bu_d)$
and $B=\lambda(\x,\bu_d)^T$. The following then applies to give a
closed-form solution to EIBV calculation in this case.

\begin{propo}
	\label{propo2}
	Let $p, q, h \geq 1$, $a \in \R^p$, $B \in \R^{p\times q}$,
	and $\covN, \covV$ be two covariance matrices in
	$\R^{p\times p}$ and $\R^{q\times q}$, respectively.
	Then, for $V \sim \mathcal{N}_{q}(0_q, \covV)$,
	\begin{equation*}
	\mathbb{E}\left[ \varPhi_{p}\left( \ba + BV; 0_p,\covN \right)^h \right]
	=
	\varPhi_{ph}
	\left(
	\ba
	;0_{ph},
	\Sigma
	\right),
	\end{equation*}
	where the vector $\ba \in \R^{p h}$ is defined as 
	$\ba := 1_h\otimes a = 
	\left(a, \dots , a
	\right)'$
	and the $p h\times p h$ covariance matrix is given by
	$\Sigma := 
	1_h 1_h'\otimes B\covV B' + I_h\otimes \covN$.
\end{propo}
The proof of Proposition \ref{propo2} is provided in the Appendix.

In block-wise representation, the following relation holds for the
covariance in Proposition \ref{propo2}: 
\begin{align*}
	\Sigma&=
	\begin{pmatrix}
	\covN & &\\
	& \ddots &\\
	&   & \covN
	\end{pmatrix}
	+
	\begin{pmatrix}
	B\covV B' & \dots & B\covV  B'\\
	\vdots & & \vdots\\
	B\covV B' & \dots & B\covV B'\\
	\end{pmatrix}.
\end{align*}
The calculation of the multivariate Gaussian cumulative distribution
is done effectively using code such as that of
\cite{genz2009computation}.

@@

\subsection{An Example of EIBV using Temperature and Salinity}
\label{Sec:UnivarEx}

@@

Table \ref{tab:sim_rhoab} (bottom two rows) shows results of expected
Bernoulli variance calculations. This is presented for a design
gathering both data types, and for a design with
temperature measurements alone. Both data are modeled by $(Y_1,Y_2)^t=(Z_1,Z_2)^t+N(0,0.5^2I_2)$, while
$Y_2=Z_2+N(0,0.5^2)$ when only temperature is measured.  For this
illustration, Table \ref{tab:sim_rhoab} shows that the expected
Bernoulli variance gets lower with larger standard deviations
$\sigma_1$ and $\sigma_2$ (right columns). The reduction of Bernoulli
variance is largest for the cases with high correlation
$\gamma$. Albeit smaller, there is also uncertainty reduction when
only temperature is measured (bottom row), especially when temperature
and salinity are highly correlated. When correlation is low
($\gamma=0.2$), there is little information about salinity in the
temperature data, and therefore less uncertainty reduction. In an
application with fresh cold water from a river source, the temperature
and salinity variables will not only be interdependent, but will also
likely show dependence in the spatial dimension. This in turn will
impact the design criteria when we evaluate the information measure by
integrating over several locations (Section \ref{sec:simulations}).

@@

An adaptive AUV survey is split into many stages. The selection on where to sample is made with restriction to the triangular waypoint graph and defined by its nearest grid nodes.

Let $d^{j,s}$ denote design
number $j$ at stage $s$ of a sequential survey. If this design is
selected, data $\by_{d^{j,s}}$ will be gathered. 
We further let
$\mathcal{Y}^{s-1} =\{(d^{j,1}, \by_{d^{j,1}}),\ldots, (d^{j,s}, \by_{d^{j,s}}) \}$ denote
the set of all data gathered until stage $s-1$ according to selected design $d^{j,1},\ldots,d^{j,s}$. 
This defines the AUV position at the stage $s$. 
The data further defines the conditional GP  with mean $\bbm^{s-1}(\x)=E(\gp[\x]|\mathcal{Y}^{s-1})$ from
Eq. (\ref{eq:cokrig})
and (cross) covariance matrix $\Sigma^{s-1}=\mbox{Var}(\gp[\x]|\mathcal{Y}^{s-1})$ from Eq. (\ref{eq:krig:cov}).

@@

The simplest heuristic for adaptive sampling is to choose the next
sampling location based on current probabilities. One reduces the largest levels of uncertainty by selecting the design $j$ with $P(Z(\x_j) \in T)$ closest to $0.5$.
For our situation with temperature and salinity variables, this {\it{naive}} strategy selects the stage $s$ sampling design according to:
\begin{eqnarray}\label{critNaive}
    d^{*,s} &=& \mbox{argmin}_{j} |P(Z_1(\x_j) \leq t_1, Z_2(\x_j) \leq t_2 | \mathcal{Y}^{s-1})-0.5|, \nonumber
\end{eqnarray}
where $j$ is indexing all neighboring locations on the waypoint graph.

@@

Based on the currently available data $\mathcal{Y}^{s-1}$, we
fit an updated GP model. Focusing on the EIBV criterion, at the next stage $s$, the selected design is:
\begin{eqnarray}\label{critSEQ}
    d^{*,s} &=& \mbox{argmin}_{j} \left\{ \int_{\mathcal{M}} E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left( p(\x;\mathcal{Y}^{j,s})\left( 1-p(\x;\mathcal{Y}^{j,s})\right) \right) d\mes(u), \right\} \\
    p(\x;\mathcal{Y}^{j,s}) &=& P(Z_u \in T |\by_{d^{j,s}},\mathcal{Y}^{s-1}). \nonumber
\end{eqnarray}

Note that this strategy gives a sequential conditional version of
expression (\ref{eq:eibv}). Now $\mathcal{Y}^{s-1}$ is available, and
the expectation is conditional to this data. A similar closed-form calculation
for EIBV is hence applicable in Eq. (\ref{critSEQ}) using the
updated GP model from stage $s-1$. Once the data is collected for the
best design at stage $s$, the GP model is updated again. The mean $\bbm^{s}$ and
covariance matrix $\Sigma^{s}$ are used to compute the next design at
stage $s+1$, and so on.

@@

We now extend the myopic strategy to a look-ahead strategy, which is
optimal when one can gather only two more stages of measurements. In
addition to the next stage of measurements $\by^{j,s}$, this
look-ahead strategy, anticipates the subsequent design indexed $j_2$ with
data $\by^{j_2,s+1}$ when choosing the current design $d^{j,s}$.  The
selected design is:
\begin{eqnarray}\label{critLA}
    d^{*,s} &=& \mbox{argmin}_{j} \left\{ E_{\by_{d^{j,s}}|\mathcal{Y}^{s-1}} \left[ \mbox{argmin}_{j_2}  V_{j_2} \right] \right\}, \\
V_{j_2} & = & \int_{\mathcal{M}} E_{\by_{d^{j_2,s+1}}|\mathcal{Y}^{j,s}} \left( p(\x;\mathcal{Y}^{j,s+1})\left( 1-p(\x;\mathcal{Y}^{j,s+1})\right) \right) d\mes(u), \nonumber \\
    p(\x;\mathcal{Y}^{j,s+1}) &=& P(Z(\x) \in T |\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}). \nonumber
\end{eqnarray}
Here, $\mathcal{Y}^{j,s}=\{\by_{d^{j,s}},\mathcal{Y}^{s-1}\}$ and
$\mathcal{Y}^{j,s+1}=\{\by_{d^{j_2,s+1}},\mathcal{Y}^{j,s}\}$ represent
the sets of data variables, and the expectations are calculated with respect to
the conditional densities given data at that stage, meaning we have two levels of intermixed minimizations and integrals.
We solve the
first expectation by Monte Carlo sampling of data $\by_{d^{j,s}}$ from its
conditional distribution. For each of these data samples, the second expectation
is solved using the closed-form expressions for EIBV from
Section \ref{sec:ESEP}.

Even though the strategy looks at two sampling stages, it is only used to find
the current best design. When data are collected, the GP model is
updated, and the mean $\bbm^{s}$ and covariance matrix $\Sigma^{s}$ are
used to compute the next design at stage $s+1$, now anticipating what
stage $s+2$ could offer, and so on.
This look-ahead approach is much more computationally demanding than
the myopic strategy, so for practical implementation, we prune paths
in the evaluation of Eq. \eqref{critLA}. This means that we use the myopic
strategy to rank the three best designs following the first stage alone, and
for each of these preferred designs, we undertake the look-ahead calculations.

@@

      \caption{Simulation results for the final mean explained variance $\bR^{2}=100*(1-(\Sigma_{posterior}/\Sigma_{initial}))$).}
